### 2013-02-11 patchy-selection-paper.tex
WorkingWiki importation

\subsection{Length of the hitchhiking haplotype}

@@

Here is a simplified picture of the process of adaptation by migration.
There is a constant outflux of migrants from the patch where the allele is already established; 
any descendants of these carrying the allele are locally maladapted, so are mostly doomed to extinction.
However, eventually the descendants of one lucky migrant will reach the new patch and colonize it.
(If the new patch is at some distance, it will almost certainly \plr{a.c.} be colonized first by only one family.)
We will estimate the total outflux of migrants, and assume that their fates are independent of each other,
so that we need only calculate the probability that offspring carrying the focal allele of a given migrant colonize the new patch,
and the distribution of the transit time, given that this occurs.
In doing so, we only follow the offspring of the migrant that carry the focal allele,
and approximate these as a branching process, 
i.e.\ assume that the number of offspring each produces is independent of each other.

>>>>

\section{Applications} 
Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX. These mice live in the desert? through the American SouthWest and into Mexico. In the majority of their range they live on the light colored rocks, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}. However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation. On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}. This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed. These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate. The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel. However, there is some evidence \citep{Hoekstra:05} that some close populations may have adapted by the spread of another dark allele by migration.

As can be seen from above a key parameter is $\sigma/\sqrt{s}$. \citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites. On the lava flow the dark MC1R allele is at $86\%$ frequency. At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency. At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass. While at the light substrate O'Neill site, which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.

I estimate that a characteristic length of 2.8km gives the frequency dropping from 50\% to ~1\% in this distance. So this doesn't seem so off. A characteristic length that is 1.4x gives 2.8\% from 50\% in 12km, which may be a little more accurate.

2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele. I estimate that we need a sigma/sqrt(s) of ~28km (i.e. ten times the characteristic length) to give a drop from 50\% to 34\%.

[Obviously the cline isn't decaying from 50\%, the frequency of the light allele on the lava to be ~14%, we could put this in more carefully]
So it seems to me that a sigma/sqrt(s) of 3-30km seems to encompass a range of interesting parameters, and can be justified without too much trouble.

Using the dispersal distance of 0.28km [from marker recapture] and an s=0.01 [from paper] gives us sigma/sqrt(s)=2.8km

%%%%%%%%%
\subsection{Length of the hitchhiking haplotype}

@@

Here is a simplified picture of the process of adaptation by migration.
There is a constant outflux of migrants from the patch where the allele is already established; 
any descendants of these carrying the allele are locally maladapted, so are mostly doomed to extinction.
(We do not require that migration be blind to the environment --
although adaptive migration would reduce the outflux of migrants --
but do assume that the focal allele does not affect migration once outside the patch.)
\plr{or maybe we should discuss this later?}
However, eventually the descendants of one lucky migrant will reach the new patch and colonize it.
(If the new patch is at some distance, it will almost certainly \plr{a.c.} be colonized first by only one family.)
We will estimate the total outflux of migrants, and assume that their fates are independent of each other,
so that we need only calculate the probability that offspring carrying the focal allele of a given migrant colonize the new patch,
and the distribution of the transit time, given that this occurs.
In doing so, we only follow the offspring of the migrant that carry the focal allele,
and approximate these as a branching process, 
i.e.\ assume that the number of offspring each produces is independent of each other.
